---
title: "FSAE Team Study"
output: html_notebook
---
Formula SAE competition at Michigan International Speedway is one of the largest and 
most competitive of the 20, international student Formula car competitions every year. 
The competition promotescareers in the research, design, manufacturing, marketing, and management of automotive and engineering careers in general. 

###Goal

look at the change in scores over the past 5 years. 
Are there any general trends year to year?  
What are the most important events in determining a team's rank? 
How many clusters of teams exist?   


```{r}
library(tidyverse)

res2017 <- read.csv("fsae_mi_2017.csv")
res2017$year <- 2017
res2016 <- read.csv("fsae_mi_2016.csv")
res2016$year <- 2016
res2015 <- read.csv("fsae_mi_2015.csv")
res2015$year <- 2015
res2014 <- read.csv("fsae_mi_2014.csv")
res2014$year <- 2014
res2013 <- read.csv("fsae_mi_2013.csv")
res2013$year <- 2013

restdl <- do.call("rbind", list(res2017,res2016,res2015,res2014,res2013))

pt <- ggplot(restdl) + 
  geom_point(aes(x=year, y = Total))
pt
```
### Observations 
1) data appears to be clustered 
2) a lot of point difference at the top points compared to the rest 
3) on average, increasing total point values awarded, even over these 5 years. 


```{r}
gatherpairs <- function(data, ..., xkey = '.xkey', xvalue = '.xvalue', ykey = '.ykey', yvalue = '.yvalue', na.rm = FALSE, convert = FALSE, factor_key = FALSE) {
  vars <- quos(...)
  xkey <- enquo(xkey)
  xvalue <- enquo(xvalue)
  ykey <- enquo(ykey)
  yvalue <- enquo(yvalue)
  
  data %>% {
    cbind(gather(., key = !!xkey, value = !!xvalue, !!!vars, na.rm = na.rm, convert = convert, factor_key = factor_key), dplyr::select(., !!!vars))
  }   %>% gather(., key = !!ykey, value = !!yvalue, !!!vars, na.rm = na.rm, convert = convert, factor_key = factor_key)
}
```

## Dynamic Events 
```{r}
restdl %>% gatherpairs( Accel, Skid_pad, Autocross, Endurance, Total) %>%
  ggplot(aes(x = .xvalue, y = .yvalue, color = Place), alpha = 0.2) +
  geom_point() +
  facet_wrap(.xkey ~ .ykey, scales = 'free')
```
### Takeaways 
1) your Endurance score is dirrectly correlated with your placement
2) Autocross and Endurance are highly correlated (when the cars complete endurance)
3) Accel shows diminishing gains but is still correlated with total points
4) skidpad is a good early indicator of who will do well in autocross and endurance 

## Static Events 
```{r}
restdl %>% gatherpairs( Cost, Presentation, Design, Efficiency, Total) %>%
  ggplot(aes(x = .xvalue, y = .yvalue, color = Place), alpha = 0.2) +
  geom_point() +
  facet_wrap(.xkey ~ .ykey, scales = 'free')
```
### Takeaways 
1) don't start focusing on Design until you are sure your car is over 600pts  
2) anyone can win the presentation event, its free points if you dedicate the time   
3) Efficiency is volitile, go hunting here for points as a last resort  
4) the faster cars tend to run more expensive, it could be a way to get an edge  


## look more closely year to year 
```{r}
normaliz <- function(value) {
  (value - min(value)) / (max(value) - min(value))
}


rt17 <- data.frame(Team = res2017$Team, year = res2017$year,  sapply(res2017[4:13], function(x) normaliz(x)))
rt16 <- data.frame(Team = res2016$Team, year = res2016$year, sapply(res2016[4:13], function(x) normaliz(x)))
rt15 <- data.frame(Team = res2015$Team, year = res2015$year, sapply(res2015[4:13], function(x) normaliz(x)))
rt14 <- data.frame(Team = res2014$Team, year = res2014$year, sapply(res2014[4:13], function(x) normaliz(x)))
rt13 <- data.frame(Team = res2013$Team, year = res2013$year, sapply(res2013[4:13], function(x) normaliz(x)))

restdl <- do.call("rbind", list(rt17, rt16, rt15, rt14, rt13))

pt <- ggplot(restdl) + 
  geom_point(aes(x=year, y = Total))
pt
```
it looks mirrored almost, 2013 and 2017 were not really competitive and 2015 was really competitive. 
its more competitive with smaller large gaps between teams, in 2016/17 it is broken down into the first 
car, the next 3, the rest of the top 10, then everyone else. However, this is just looking at the general competition, not the 

### Team Trends 
```{r}
# look at only teams that are in dataset for all 5 years 
# plot each team's trend 
# 
sbst <- table(restdl$Team)

sbst <- ifelse(sbst<5, FALSE, TRUE)

trnds <- data.frame(matrix(vector(), 0, 12))
names(trnds) <- names(restdl)
for(i in 1:nrow(restdl)){
  id <- restdl$Team[i]
  
    if(sbst[which(names(sbst)==id)] == TRUE){
      trnds <- rbind(trnds, restdl[i,])
      
  }
}

nrow(trnds)/5
trnds$Team <- as.factor(as.character((trnds$Team)))
trnds[is.na(trnds)] <- 0 
tmanes <- names(table(trnds$Team))
```

```{r}
trnds.sb <- trnds[which(trnds$Team %in% tmanes[1:7]),]
pt <- ggplot(trnds.sb) + 
  geom_point(aes(x = year, y = Total, col = Team)) + 
  geom_line(aes(x = year, y = Total, col = Team)) + 
  xlim(2013, 2017)
pt
```


```{r}
trnds.sb <- trnds[which(trnds$Team %in% tmanes[8:14]),]
pt <- ggplot(trnds.sb) + 
  geom_point(aes(x = year, y = Total, col = Team)) + 
  geom_line(aes(x = year, y = Total, col = Team)) + 
  xlim(2013, 2017)
pt
```


```{r}
trnds.sb <- trnds[which(trnds$Team %in% tmanes[15:21]),]
pt <- ggplot(trnds.sb) + 
  geom_point(aes(x = year, y = Total, col = Team)) + 
  geom_line(aes(x = year, y = Total, col = Team)) + 
  xlim(2013, 2017)
pt
```

```{r}
trnds.sb <- trnds[which(trnds$Team %in% tmanes[22:27]),]
pt <- ggplot(trnds.sb) + 
  geom_point(aes(x = year, y = Total, col = Team)) + 
  geom_line(aes(x = year, y = Total, col = Team)) + 
  xlim(2013, 2017)
pt
```

### cluster groups on pca
The Team name and year of competition are not included in the fit. Team name is ultimately what we want to know from this clustering so including it in the cluster will produce weird fits. The year of competition should have been accounted for in the normalization process (everything rescaled on [0,1] with only teams that competed in every year included in the dataset) so it is safe to not include it in this fit. 
```{r}

pca <- prcomp(trnds[3:12])
print(pca)
plot(pca)
```
Based on the plot, the top 3 principal components will be used 
PC1 = Endurance, Efficiency, Total Points 
PC2 = Acceleration, Skid_pad 
PC3 = Acceleration, Presentation

### Cluster Teams
plot data on the first 2 principal components 
```{r}
topteams <- kmeans(pca$x[,1:3], centers = 4, iter.max = 100, nstart = 1000,  algorithm = "Hartigan-Wong")
trnds$cluster <- as.factor(topteams$cluster)

pt <- ggplot(trnds) + 
  geom_point(aes(x = pca$x[,2], y = pca$x[,1], col = cluster))
pt
```
Looking at the data on the previous axes we have been looking at. 
```{R}
pt <- ggplot(trnds) + 
  geom_point(aes(x = year, y = Total, col = cluster))
pt
```

The 27 teams we were investigating and their most often associated cluster: (alphebetical by team name) 

```{r}
clanes <- vector(length = 27)
for(i in 1:length(tmanes)){
  results <- as.numeric(trnds[which(trnds$Team %in% tmanes[i]),]$cluster)
  clanes[i] <- as.numeric(names(which.max(table(results))))
  
}

print(data.frame(Team = tmanes,cluster = clanes))
```
### Predict All Team Clusters 
```{r}
library(clue)
pca.all <- predict(pca, newdata = restdl[3:12])
teamcl <- cl_predict(topteams, pca.all[,1:3])
tmanes.n <- names(table(restdl$Team))
restdl$cluster <- as.factor(teamcl)


for(i in 1:length(tmanes.n)){
  results <- as.numeric(restdl[which(restdl$Team %in% tmanes.n[i]),]$cluster)
  clanes[i] <- as.numeric(names(which.max(table(results))))
  
}
print(data.frame(team = tmanes.n,cluster = clanes))
```




