---
title: "FSAE Time-Trial Analysis "
Author: "Nathan Krumholz" 
output: html_document
---

Have you ever seen a video comparing Olympic sports from the Seventies to modern day? 
They difference between the two is striking, what is required to even make the cut nowadays completely eclipses what would have won their respective events before. 
This phenomenon didn?t just occur overnight; instead there was creep, an incremental increase in difficulty year to year perpetuated by a number of different factors. 
These factors are not of concern in this paper, rather what the rate of creep experienced in the last few years of FSAE. 
The goal of this analysis is to determine the rate of time decreases in event times that a FSAE team needs to acquire just to remain competitive. 

There are multiple ways to sample the data for analysis, here is the method used in this analysis: 

Seven teams from the top performing cluster in the previous analysis were randomly sampled, with a preference for teams with more years.
Taking more data points would have taken more time and any less might not have been enough to determine any trends. 
Taking just the top performing teams of every event is not a good idea, a team might build a car specifically to win acceleration but not to win any other event.
A team like this is not playing the same game as the other teams to which their creep factor may be unrelated to the competition as a whole. 
Taking the top overall performing teams from every year is another method (and would be a good extension to this analysis). 
However, you lose out on not necessarily following the same teams? year to year, thus not seeing how the same vehicle gets refined each year to achieve the creep.  

```{r}
library(tidyverse)

# Load data
setwd("E:/ME/PostGrad/FSAE analysis/Points Analysis/Points Analysis")
X <- read.csv("Top 10 trends.csv")
tten <- subset(X, X$Year < 2018)
tten8 <- subset(X, X$Year == 2018)
```

### Placements
```{r, echo = FALSE}
## Placements
ptp <- ggplot(X) +
    geom_point(aes(x = as.factor(Year), y = Place, col = Team)) +
    geom_boxplot(aes(x = as.factor(Year), y = Place), alpha = 0.25) +
    ylab("Final Placement")+ xlab("Year") + ggtitle("Placement Trends of Sampled Teams")
ptp
```
Based on the plot and the fit coefficients there does not appear to much change in the mean 
placement from year to year (appart from 2013) . This helps to ensure that the analysis is comparing 
similar data. 

### Acceleration
```{r, echo=FALSE}
## Acceleration Event 
tnd_A <- lm(tten$Accel ~ tten$Year)
tnd_A
# intercept -28.483
# trends 0.0163
```

```{r}
pta <- ggplot(X) +
    geom_point(aes(x = Year, y = Accel)) +
    geom_abline(intercept = tnd_A$coefficients[1], slope = tnd_A$coefficients[2]) +
    ylab("Seconds") + ggtitle("Acceleration Event Trends")
pta
```


The trend shows a slight annual increase in Acceleration times of apporximatley 0.015 seconds per year. 
This is relatively insignificant and is most likely due to the proliferation of aerodynamics and downforce generation. 

### SkidPad
```{r, echo=FALSE}
tnd_S <- lm(tten$Skid_pad ~ tten$Year)
tnd_S
# intercept 103.78
# trend -0.0489
```

```{r}
pts <- ggplot(X) +
    geom_point(aes(x = Year, y = Skid_pad)) +
    geom_abline(intercept = tnd_S$coefficients[1], slope = tnd_S$coefficients[2]) +
    ylab("Seconds") + ggtitle("Skidpad Event Trends")
pts
```
There appears to be a significant decrease in times per year on skidpad leading to a decrease 
a quarter of a second on this event over five years.
 
### Autocross
```{r, echo=FALSE}
## Autocross Event
tnd_C <- lm(tten$Autocross ~ tten$Year)
tnd_C
# intercept 3741.59
# trend -1.83
```

```{r}
ptc <- ggplot(X) +
    geom_point(aes(x = Year, y = Autocross)) +
    geom_abline(intercept = tnd_C$coefficients[1], slope = tnd_C$coefficients[2]) +
    ylab("Seconds") + ggtitle("Autocross Event Trends")
ptc
```

This analysis shows that there is over a 1.75 second decrease in autocross laptimes per year. 
It is unclear if 2017 was just a fast autocross track or if the trend does exist or some combination 
of the two. The data could also show a four year cycle where 2013 and 2017 were fast with the other years 
slower tracks, but that is unclear from this data snapshot. 

### Endurance
```{r, echo=FALSE}
## Endurance Event 
tnd_E <- lm(tten$Endurance ~ tten$Year)
tnd_E
# intercept -8072.16
# trend 4.722
```

```{r}
pte <- ggplot(X) +
    geom_point(aes(x = Year, y = Endurance)) +
    geom_abline(intercept = tnd_E$coefficients[1], slope = tnd_E$coefficients[2]) +
    ylab("Seconds") + ggtitle("Endurance Event Trends")
pte
```
Even though the autocross event saw a decrease year to year in times the Endurance times appear to be increasing. 
This is most likely due to course design where the courses are getting longer every year despite the continuous improvements
by the teams. This window of data could show a three year cycle where 2014 and 
2017 are fast tracks, 2015 and 2018 are slow tracks, and 2013 and 2016 are relatively slow. Keep that in mind 
when comparing Endurance times in the future. 

### 2018 conclusions
```{r, echo=FALSE}
## 2018 predictions
accel <- tnd_A$coefficients[1] + tnd_A$coefficients[2] * 2018
skidpad <- tnd_S$coefficients[1] + tnd_S$coefficients[2] * 2018
autocross <- tnd_C$coefficients[1] + tnd_C$coefficients[2] * 2018
endurance <- tnd_E$coefficients[1] + tnd_E$coefficients[2] * 2018
```

```{r}
bpa <- ggplot(tten8) +
    geom_boxplot(aes(y = Accel, x = as.factor(Year))) +
    geom_point(aes(y = accel, x = as.factor(2018)), col = "red", size = 3) +
    xlab("Year") + ylab("seconds") + ggtitle("Actual vs. Prediction for Acceleration Event")
bpa
```
```{r}
bps <- ggplot(tten8) +
    geom_boxplot(aes(y = Skid_pad, x = as.factor(Year))) +
    geom_point(aes(y = skidpad, x = as.factor(2018)), col = "red", size = 3) +
    xlab("Year") + ylab("seconds") + ggtitle("Actual vs. Prediction for Skid Pad Event")
bps
```
```{r}
bpc <- ggplot(tten8) +
    geom_boxplot(aes(y = Autocross, x = as.factor(Year))) +
    geom_point(aes(y = autocross, x = as.factor(2018)), col = "red", size = 3) +
    xlab("Year") + ylab("seconds") + ggtitle("Actual vs. Prediction for Autocross Event")
bpc

```
```{r}
bpe <- ggplot(tten8) +
    geom_boxplot(aes(y = Endurance, x = as.factor(Year))) +
    geom_point(aes(y = endurance, x = as.factor(2018)), col = "red", size = 3) +
    xlab("Year") + ylab("seconds") + ggtitle("Actual vs. Prediction for Endurance Event")
bpe
```

All of our 2018 time predictions were faster than the mean of the poulation but most were still within
the lower quartile of the datapoints on all of the events. Keep in mind that it rained a lot in 2018 
and that more cars than normal (even in the utilized dataset) did not complete the endurace event. 

I would warn teams about looking at these numbers and forgetting to take weather effects into consideration. 
It is unclear if the trends exhibited in this analysis are correct due to the uncharateristic nature 
of the 2018 competition. 
